Reliable error estimation for Sobol' indices

### 2016-06-10 article.tex
changes article.tex: annesection 1 + annexe

We define by design a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1)^d$. Each column of the design refers to a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid [0,1)^{2d}$. One way to estimate the two quantities (\ref{first.order}) and (\ref{total.effect}) is via:
\begin{align}
\label{first.order.est}
\widehat{\underline{\tau}_u^2} & = \frac{1}{n} \sum \limits_{i=0}^{n-1} (f(\vx_i)-\hat{\mu})(f(\vx_{i,u}:{\vx'}_{i,-u})-\hat{\mu}),\\
\label{total.effect.est}
\widehat{\ov{$\tau$}_u^2} & = \frac{1}{2n} \sum \limits_{i=0}^{n-1} (f(\vx_i) - f({\vx'}_{i,u}:\vx_{i,-u}))^2,
\end{align}
using for $\mu$ the efficient estimator introduced in \cite{Monod, Janon}:
\[ \hat{\mu} = \frac{1}{2n} \sum \limits_{i=0}^{n-1} f(\vx_i) + f({\vx'}_{i,u}:\vx_{i,-u}). \]
The estimation of a single pair ($\underline{\tau}_u^2$, $\ov{$\tau$}_u^2$) requires $3n$ evaluations of the model $f$. Using a combinatorial formalism, Saltelli \cite[Theorem 1]{Saltelli} proposed the following strategy to reduce the estimation cost:
\begin{theorem}
\label{saltelli.theorem}
The $d+2$ designs $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i=0}^{n-1}$ constructed for $u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$ allows to estimate all first-order and all total effect Sobol' indices at a cost of $n(d+2)$ evaluations of the model.
\end{theorem}
The $d+2$ designs of Theorem \ref{saltelli.theorem} are obtained by substituting columns of $\mathcal{P}$ for columns of $\mathcal{P}'$ accordingly to subset $u$. While elegant, this approach still requires a number of model evaluations that grows linearly with respect to the input space dimension.

@@

\subsection{Iterative estimation procedures}
We propose here two iterative procedures two estimate first-order and total effect Sobol' indices. Each procedure combines quasi-Monte Carlo sampling with respectively the estimation strategy proposed by Saltelli (Theorem \ref{saltelli.theorem}) and the estimation strategy based on replicated designs. In the following, we refer to these two iterative procedures as procedure $A$ and $B$.
\bigskip

Both procedures require to iteratively construct two points sets $\mathcal{P}_\ell=\{\vx_i\}_{i=0}^{2^\ell-1}$ and $\mathcal{P}'_\ell=\{{\vx'}_i\}_{i=0}^{2^\ell-1}$ that identify as two Sobol' sequences. Background on Sobol' sequences are provided in \cite{crass}. The construction of these two points sets is carried out according to the following recursive scheme :
$$\left\lbrace \begin{array}{l}
\mathcal{P}_0= B_0 \\
\mathcal{P}_\ell= \mathcal{P}_{\ell-1} \cup B_\ell \end{array}\right. ,\qquad
\hspace*{0.5cm}
\left\lbrace \begin{array}{l}
\mathcal{P}'_0= {B'}_0 \\
\mathcal{P}'_\ell= \mathcal{P}'_{\ell-1} \cup {B'}_\ell \end{array}\right.,
$$
where at step $\ell\geq0$, $B_\ell$ and $B'_\ell$ are new sets of $2^{\ell-1}$ points added to refine $\mathcal{P}_{\ell-1}$ and $\mathcal{P}'_{\ell-1}$. In the following, $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ refer to the two Sobol' sequences constructed following the multiplicative approach presented in \cite{crass}. This approach guarantee that at each step $\ell\geq0$, $\mathcal{P}_{\ell}$ and $\mathcal{P}'_{\ell}$ are two replicated designs of order $1$.

$\mathcal{P}_{\ell}$ and $\mathcal{P}'_{\ell}$ can either be used with procedure $A$ to estimate all first-order and all total effect Sobol' indices or with procedure $B$ to estimate all first-order Sobol' indices. Algorithms \ref{procedureA} and \ref{procedureB} summarize the main steps of each iterative estimation procedure.
\begin{algorithm}[!ht]
\caption{Iterative estimation: procedure $A$}
\begin{algorithmic}[1]
\vspace*{0.2cm}
\State Set: $\ell \leftarrow 1$, $\mathcal{P}_{0} \leftarrow B_0$, ${\mathcal{P}'}_{0} \leftarrow {B'}_0$
\While {stopping criterion}
\State $\mathcal{P}_\ell \leftarrow \mathcal{P}_{\ell-1} \cup B_\ell$

\hspace*{-0.3cm} $\mathcal{P}'_\ell \leftarrow \mathcal{P}'_{\ell-1} \cup {B'}_\ell$
\For {$u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$}
\State Construct design $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i=0}^{2^{\ell}-1}$
\State Compute $\{f(\vx_i)\}_{i=0}^{2^\ell-1}$ and $\{f(\vx_{i,u},{\vx'}_{i,-u})\}_{i=0}^{2^\ell-1}$
\State Estimate $\widehat{\underline{\tau}_u^2}$ and $\widehat{\ov{$\tau$}_u^2}$ with formula (\ref{first.order.est}) and (\ref{total.effect.est})
\EndFor
\State $\ell \leftarrow \ell + 1$
\EndWhile
\State Return $\widehat{\underline{\tau}_u^2}$ and $\widehat{\ov{$\tau$}_u^2}$ for all $u \in \mathcal{D}$.
\end{algorithmic}
\label{procedureA}
\end{algorithm}
\begin{algorithm}[!ht]
\caption{Iterative estimation: procedure $B$}
\begin{algorithmic}[1]
\vspace*{0.2cm}
\State Set: $\ell \leftarrow 1$, $\mathcal{P}_{0} \leftarrow B_0$, ${\mathcal{P}'}_{0} \leftarrow {B'}_0$
\While {stopping criterion}
\State $\mathcal{P}_\ell \leftarrow \mathcal{P}_{\ell-1} \cup B_\ell$

\hspace*{-0.3cm} $\mathcal{P}'_\ell \leftarrow \mathcal{P}'_{\ell-1} \cup {B'}_\ell$
\State Compute $\{f(\vx_i)\}_{i=0}^{2^\ell-1}$ and $\{f({\vx'}_i)\}_{i=0}^{2^\ell-1}$
\For {$u \in \mathcal{D}$}
\State  $\{f({\vx'}_i)\}_{i=0}^{2^\ell-1}$ into $\{\pi_u(f({\vx'}_i))\}_{i=0}^{2^\ell-1}$
\State Estimate $\widehat{\underline{\tau}_u^2}$ with formula (\ref{first.order.est})
\EndFor
\State $\ell \leftarrow \ell + 1$
\EndWhile
\State Return $\widehat{\underline{\tau}_u^2}$ for all $u \in \mathcal{D}$.
\end{algorithmic}
\label{procedureB}
\end{algorithm}

The choice of the stopping criterion is the key point of this paper. In most iterative approaches that estimate Sobol' indices, the stopping criterion is a quantity of interest build directly from the estimates. For example, an absolute difference between estimates obtained on consecutive steps. Such stopping criteria are often hard to tweak but above all fail to guarantee any error bounds on the Sobol' indices. Our stopping criterion is presented and discussed in Section \ref{section.error}. The criterion is an error bound based on the discrete Walsh decomposition of the integral formula of the first-order and total effect Sobol' index. The formulation of this error bound makes use of the closed set property of Sobol' sequences.

\subsection{Improvements}
This section provide two improvements to ease the estimation of Sobol' indices with procedures $A$ and $B$. The first improvement focus on the use of a better estimator to evaluate small first-order Sobol' indices with procedure $A$. The second improvement focus on the choice of the $2d$ generating matrices that generate the two Sobol' sequences $\mathcal{P}_\ell$  and ${\mathcal{P}'}_\ell$ in procedure $B$.

@@

This estimator called ``Correlation 2" was introduced by Owen in \cite{Owen}. Owen discussed and highlighted the efficiency of ``Correlation 2" when estimating small closed Sobol' indices. Our aim is to show that using ``Correlation 2" to estimate small first-order Sobol' indices in procedure $A$  reduces the total number of model evaluations required while reaching the same accuracy on the estimates.

@@

Cost small - cost pas small: $n\sum_{i=1}^{s}(2-\alpha_i+1)$ $s$ number of small indices, $\alpha_i$ proportion of how many more points if not pas small
\bigskip

The choice of the $2d$ generating matrices that generate the two Sobol' sequences $\mathcal{P}_\ell$ and ${\mathcal{P}'}_\ell$ in the multiplicative approach \cite{crass}. These generating matrices are selected from the set found by Kuo and Joe in \cite{Kuo}. This set is composed of generating matrices that optimize the $2$-dimensional projection $t$-values. Elements of the set are ordered based on the best optimized $t$-values.

@@

>>>>

We define by design a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1)^d$. Each column of the design refers to a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid [0,1)^{2d}$. One way to estimate the two quantities (\ref{first.order}) and (\ref{total.effect}) is via:
\begin{align}
\label{first.order.est}
\widehat{\underline{\tau}_u^2} & = \frac{1}{n} \sum \limits_{i=0}^{n-1} (f(\vx_i)-\hat{\mu})(f(\vx_{i,u}:{\vx'}_{i,-u})-\hat{\mu}),\\
\label{total.effect.est}
\widehat{\ov{$\tau$}_u^2} & = \frac{1}{2n} \sum \limits_{i=0}^{n-1} (f(\vx_i) - f({\vx'}_{i,u}:\vx_{i,-u}))^2,
\end{align}
using for $\mu$ the efficient estimator introduced in \cite{Monod, Janon}:
\begin{equation}
 \hat{\mu} = \frac{1}{2n} \sum \limits_{i=0}^{n-1} f(\vx_i) + f(\vx_{i,u}:{\vx'}_{i,-u}).
\label{mu.est}
\end{equation}
The estimation of a single pair ($\underline{\tau}_u^2$, $\ov{$\tau$}_u^2$) requires $3n$ evaluations of the model $f$. Using a combinatorial formalism, Saltelli \cite[Theorem 1]{Saltelli} proposed the following strategy to reduce the estimation cost:
\begin{theorem}
\label{saltelli.theorem}
The $d+2$ designs $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i=0}^{n-1}$ constructed for $u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$ allows to estimate all first-order and all total effect Sobol' indices at a cost of $n(d+2)$ evaluations of the model.
\end{theorem}
The $d+2$ designs of Theorem \ref{saltelli.theorem} are obtained by substituting columns of $\mathcal{P}$ for columns of $\mathcal{P}'$ accordingly to subset $u$. While elegant, this approach still requires a number of model evaluations that grows linearly with respect to the input space dimension.

@@

\subsection{Recursive estimation procedures}
We propose here two recursive procedures two estimate first-order and total effect Sobol' indices. Each procedure combines quasi-Monte Carlo sampling with respectively the estimation strategy proposed by Saltelli (Theorem \ref{saltelli.theorem}) and the estimation strategy based on replicated designs. In the following, we refer to these two procedures as procedure $A$ and $B$.
\bigskip

Using a recursive estimation procedure requires to write a recursive formula for estimators (\ref{first.order.est}) and (\ref{total.effect.est}). For the sake of readability of the paper, the corresponding recursive formulas are detailed in equations (\ref{first.order.est.rec}) and (\ref{total.effect.est.rec}) in the Annexe.

Procedures $A$ and $B$ also require to iteratively construct two designs $\mathcal{P}_\ell=\{\vx_i\}_{i=0}^{2^\ell-1}$ and $\mathcal{P}'_\ell=\{{\vx'}_i\}_{i=0}^{2^\ell-1}$. These two designs are two Sobol' sequences. Background on Sobol' sequences are provided in \cite{crass}. The construction of $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ is carried out according to the following recursive scheme :
\begin{equation}
\left\lbrace \begin{array}{l}
\mathcal{P}_0= B_0 \\
\mathcal{P}_\ell= \mathcal{P}_{\ell-1} \cup B_\ell \end{array}\right. ,\qquad
\hspace*{0.5cm}
\left\lbrace \begin{array}{l}
\mathcal{P}'_0= {B'}_0 \\
\mathcal{P}'_\ell= \mathcal{P}'_{\ell-1} \cup {B'}_\ell \end{array}\right.,
\label{rec.scheme}
\end{equation}
where at step $\ell\geq0$, $B_\ell$ and $B'_\ell$ are new sets of $2^{\ell-1}$ points added to refine $\mathcal{P}_{\ell-1}$ and $\mathcal{P}'_{\ell-1}$. In the following, $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ refer to the two Sobol' sequences constructed following the multiplicative approach presented in \cite{crass}. This approach guarantee that at each step $\ell\geq0$, $\mathcal{P}_{\ell}$ and $\mathcal{P}'_{\ell}$ are two replicated designs of order $1$.

$\mathcal{P}_{\ell}$ and $\mathcal{P}'_{\ell}$ can either be used with procedure $A$ to estimate all first-order and all total effect Sobol' indices or with procedure $B$ to estimate all first-order Sobol' indices. Algorithms \ref{procedureA} and \ref{procedureB} summarize the main steps of each recursive estimation procedure.
\begin{algorithm}[!ht]
\caption{Procedure $A$}
\begin{algorithmic}[1]
\vspace*{0.2cm}
\State Set: $\ell \leftarrow 1$, $\mathcal{P}_{0} \leftarrow B_0$, ${\mathcal{P}'}_{0} \leftarrow {B'}_0$
\While {stopping criterion}
\State $\mathcal{P}_\ell \leftarrow \mathcal{P}_{\ell-1} \cup B_\ell$

\hspace*{-0.3cm} $\mathcal{P}'_\ell \leftarrow \mathcal{P}'_{\ell-1} \cup {B'}_\ell$
\For {$u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$}
\State Construct design $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i=0}^{2^{\ell}-1}$
\State Compute $\{f(\vx_i)\}_{i=0}^{2^\ell-1}$ and $\{f(\vx_{i,u},{\vx'}_{i,-u})\}_{i=0}^{2^\ell-1}$
\State Estimate $\widehat{\underline{\tau}_u^2}^{(\ell)}$ and $\widehat{\ov{$\tau$}_u^2}^{(\ell)}$ with formula (\ref{first.order.est}) and (\ref{total.effect.est})
\EndFor
\State $\ell \leftarrow \ell + 1$
\EndWhile
\State Return $\widehat{\underline{\tau}_u^2}^{(\ell)}$ and $\widehat{\ov{$\tau$}_u^2}^{(\ell)}$ for all $u \in \mathcal{D}$.
\end{algorithmic}
\label{procedureA}
\end{algorithm}
\begin{algorithm}[!ht]
\caption{Procedure $B$}
\begin{algorithmic}[1]
\vspace*{0.2cm}
\State Set: $\ell \leftarrow 1$, $\mathcal{P}_{0} \leftarrow B_0$, ${\mathcal{P}'}_{0} \leftarrow {B'}_0$
\While {stopping criterion}
\State $\mathcal{P}_\ell \leftarrow \mathcal{P}_{\ell-1} \cup B_\ell$

\hspace*{-0.3cm} $\mathcal{P}'_\ell \leftarrow \mathcal{P}'_{\ell-1} \cup {B'}_\ell$
\State Compute $\{f(\vx_i)\}_{i=0}^{2^\ell-1}$ and $\{f({\vx'}_i)\}_{i=0}^{2^\ell-1}$
\For {$u \in \mathcal{D}$}
\State  $\{f({\vx'}_i)\}_{i=0}^{2^\ell-1}$ into $\{\pi_u(f({\vx'}_i))\}_{i=0}^{2^\ell-1}$
\State Estimate $\widehat{\underline{\tau}_u^2}^{(\ell)}$ with formula (\ref{first.order.est})
\EndFor
\State $\ell \leftarrow \ell + 1$
\EndWhile
\State Return $\widehat{\underline{\tau}_u^2}^{(\ell)}$ for all $u \in \mathcal{D}$.
\end{algorithmic}
\label{procedureB}
\end{algorithm}

The choice of the stopping criterion is the key point of this paper. In most recursive approaches that estimate Sobol' indices, the stopping criterion is a quantity of interest build directly from the estimates. For example, an absolute difference between estimates computed on consecutive steps. Such stopping criteria are often hard to tweak but above all fail to guarantee any error bounds on the estimation. Our stopping criterion is presented and discussed in Section \ref{section.error}. The criterion is an error bound based on the discrete Walsh decomposition of the integral formulas of first-order and total effect Sobol' indices.

\subsection{Improvements}
This section discusses two improvements to ease the estimation of Sobol' indices with procedures $A$ and $B$.

@@

The first improvement focus on the use of a better estimator to evaluate small first-order Sobol' indices with procedure $A$. This estimator is called ``Correlation 2" and was introduced by Owen in \cite{Owen}. Owen discussed and highlighted the efficiency of ``Correlation 2" when estimating small closed Sobol' indices.

Our aim is to show that using ``Correlation 2" in procedure $A$ may reduces the total number of model evaluations required while reaching the same accuracy on the estimates.

@@

Cost small - cost pas small: $n\sum_{i=1}^{s}(2-\alpha_i+1)$ $s$ number of small indices, $\alpha_i$ proportion of how many more points if not pas small

The second improvement focus on the choice of the $2d$ generating matrices that generate the two Sobol' sequences $\mathcal{P}_\ell$  and ${\mathcal{P}'}_\ell$ in procedure $B$.

In the multiplicative approach \cite{crass}, these $2d$ generating matrices are selected as the first $2d$ elements of the set found by Kuo and Joe in \cite{Kuo}. This set is composed of generating matrices that optimize the $2$-dimensional projection $t$-values. For procedure $A$ this choice of these $2d$ generating matrices is logical. Each of the $d+2$ designs constructed following Theorem \ref{saltelli.theorem} is a Sobol' sequence generated by a subset of $d$ matrices among the initial $2d$ ones. Therefore, $2$-dimensional projection $t$-values are not changed.

That is not the case for procedure $B$. (introduire les matrices $C_j$ et ecrire les relations). The permutations applied in step $6$ of Algorithm \ref{procedureB} modify the $2$-dimensional projection $t$-values of the underlying designs.

@@

\newpage
\section{Annexe: recursive formulas}

Let $\mathcal{P}_\ell=\{\vx_i\}_{i=0}^{2^\ell-1}$ and $\mathcal{P}'_\ell=\{{\vx'}_i\}_{i=0}^{2^\ell-1}$ be the two points sets constructed with the recursive scheme (\ref{rec.scheme}).

\subsection{Recursive formula for $\hat{\mu}$}

The quantity $\hat{\mu}$ defined by formula (\ref{mu.est}) can be recursively estimated by:
\[\left \lbrace\begin{array}{l}
\hat{\mu}^{(0)} = \dfrac{1}{2} \Big{(} f(\vx_0)+f(\vx_{0,u}:{\vx'}_{0,-u}) \Big{)} \\
\\
\hat{\mu}^{(\ell+1)} = \dfrac{1}{2} \Big{(} \hat{\mu}^{(\ell)} + \dfrac{1}{2^{\ell+1}} \sum \limits_{i=2^{\ell}}^{2^{\ell+1}-1} f(\vx_i) + f(\vx_{i,u}:{\vx'}_{i,-u}) \Big{)}
\end{array}\right.\]
\begin{proof}
By developing $\hat{\mu}^{(\ell+1)}$:
\begin{align*}
\hat{\mu}^{(\ell+1)} &= \dfrac{1}{2} \dfrac{1}{2^{\ell+1}} \sum \limits_{i=0}^{2^{\ell+1}-1} f(\vx_i) + f(\vx_{i,u}:{\vx'}_{i,-u}) \\
\hat{\mu}^{(\ell+1)} &= \dfrac{1}{2} \dfrac{1}{2^{\ell+1}} \Big{(} \sum \limits_{i=0}^{2^{\ell}-1} f(\vx_i) + f(\vx_{i,u}:{\vx'}_{i,-u}) + \sum \limits_{i=2^{\ell}}^{2^{\ell+1}-1} f(\vx_i) + f(\vx_{i,u}:{\vx'}_{i,-u}) \Big{)}\\
\hat{\mu}^{(\ell+1)} &= \dfrac{1}{2} \dfrac{1}{2^{\ell+1}} \Big{(} 2 \times 2^{\ell}\hat{\mu}^{(\ell)} + \sum \limits_{i=2^{\ell}}^{2^{\ell+1}-1} f(\vx_i) + f(\vx_{i,u}:{\vx'}_{i,-u}) \Big{)}\\
\end{align*}
\end{proof}

\subsection{Recursive formula for $\widehat{\underline{\tau}_u^2}$}

The quantity $\widehat{\underline{\tau}_u^2}$ defined by formula (\ref{first.order.est}) can be recursively estimated by:
\begin{equation}
\left\lbrace \begin{array}{l}
\widehat{\underline{\tau}_u^2}^{(0)}= (f(\vx_0)-\hat{\mu}^{(0)})(f(\vx_{0,u}:{\vx'}_{0,-u})-\hat{\mu}^{(0)}) \\
\\
\widehat{\underline{\tau}_u^2}^{(\ell+1)} =  \dfrac{1}{2} \widehat{\underline{\tau}_u^2}^{(\ell)} - \dfrac{1}{2} (\hat{\mu}^{(\ell)})^2 - (\hat{\mu}^{(\ell+1)})^2 + \dfrac{1}{2^{\ell+1}} \sum \limits_{i=2^{\ell}+1}^{2^{\ell+1}-1} f(\vx_i) f(\vx_{i,u}:{\vx'}_{i,-u})
\end{array}\right.
\label{first.order.est.rec}
\end{equation}
\begin{proof}
By developing $\widehat{\underline{\tau}_u^2}^{(\ell+1)}$:
\begin{align*}
\widehat{\underline{\tau}_u^2}^{(\ell+1)} &= \frac{1}{2^{\ell+1}} \sum \limits_{i=0}^{2^{\ell+1}-1} (f(\vx_i)-\hat{\mu}^{(\ell+1)})(f(\vx_{i,u}:{\vx'}_{i,-u})-\hat{\mu}^{(\ell+1)}) \\
\widehat{\underline{\tau}_u^2}^{(\ell+1)} &= \frac{1}{2^{\ell+1}} \sum \limits_{i=0}^{2^{\ell+1}-1} f(\vx_i)f(\vx_{i,u}:{\vx'}_{i,-u})-(\hat{\mu}^{(\ell+1)})^2 \\
\widehat{\underline{\tau}_u^2}^{(\ell+1)} &= \frac{1}{2^{\ell+1}} \Big{(}\sum \limits_{i=0}^{2^{\ell}-1} f(\vx_i)f(\vx_{i,u}:{\vx'}_{i,-u}) + \sum \limits_{i=2^{\ell}}^{2^{\ell+1}-1} f(\vx_i)f(\vx_{i,u}:{\vx'}_{i,-u})\Big{)} -(\hat{\mu}^{(\ell+1)})^2 \\
\widehat{\underline{\tau}_u^2}^{(\ell+1)} &= \dfrac{1}{2}\Big{(}\widehat{\underline{\tau}_u^2}^{(\ell)}-(\hat{\mu}^{(\ell)})^2\Big{)} + \frac{1}{2^{\ell+1}} \sum \limits_{i=2^{\ell}}^{2^{\ell+1}-1} f(\vx_i)f(\vx_{i,u}:{\vx'}_{i,-u})-(\hat{\mu}^{(\ell+1)})^2 \\
\end{align*}
\end{proof}

\subsection{Recursive formula for} 
%$\widehat{\ov{$\tau$}_u^2}$}

The quantity $\widehat{\underline{\tau}_u^2}$ defined by formula (\ref{total.effect.est}) can be recursively estimated by:
\begin{equation}
\left\lbrace \begin{array}{l}
\widehat{\ov{$\tau$}_u^2}^{(0)}= \dfrac{1}{2} \Big{(} f(\vx_0)-(f({\vx'}_{0,u}:\vx_{0,-u}) \Big{)}^2 \\
\\
\widehat{\ov{$\tau$}_u^2}^{(\ell+1)} = \dfrac{1}{2} \widehat{\ov{$\tau$}_u^2}^{(\ell)} + \dfrac{1}{2^{\ell+1}} \sum \limits_{i=2^{\ell}+1}^{2^{\ell+1}-1} \Big{(}f(\vx_i)- f({\vx'}_{i,u}:\vx_{i,-u})\Big{)}^2 
\end{array}\right.
\label{total.effect.est.rec}
\end{equation}
\begin{proof}
Straightforward by developing $\widehat{\ov{$\tau$}_u^2}^{(\ell+1)}$.
\end{proof}
